---
title: "Knapsack package"
author: "Henrik Karlsson and Milda Pocevicuite"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Knapsack package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Data used
The following example data is being used for all knapsack solutions in this package. 

Our package depends on dplyr.

```{r data, warning = FALSE, comment= NA, message=FALSE}
library(dplyr)

set.seed(42)
n <- 2000
knapsack_objects <-
    data.frame(
        w=sample(1:4000, size = n, replace = TRUE),
        v=runif(n = n, 0, 10000)
    )
```

# Brute Force Search


# Dynamic Programming


# Greedy Heuristic

The greedy_heuristic function takes two arguments, **x** which is the data frame containing weight and value and **w** which is the total maximum allowed weight of the function. The function computes the density $density = \frac{v}{w}$, sorts the data in in descending density order and pick out the objects with higest density within the the allowed maximum weight, w.

```{r, greedy_heuristic}
greedy_heuristic <- function(x = knapsack_objects, w){
    stopifnot(is.numeric(w))
    
    x <- x %>%
        mutate(element = seq_along((x[[1]])),
               density = v/w) %>% 
        arrange(desc(density))
    
    # Start values
    sum_weight <- 0
    sum_value <- 0
    weight_check <- 0
    selected_elements <- vector()
    i <- 1
    
    while (weight_check <= w) {
        sum_weight <- sum_weight + x[i, "w"]
        sum_value <- sum_value + x[i, "v"]
        
        selected_elements[i] <- x[i, "element"]
        
        weight_check <- sum_weight + x[i + 1, "w"]
        i <- i + 1
    }
    
    return(list(value = round(sum_value, 0),
                sum_weight = sum_weight,
                selected_elements = selected_elements))
}
```

### How much time does it take to run the algorithm?
```{r, gh_time}
n_gh <- 1000000


start <- Sys.time()
greedy_heuristic(x = knapsack_objects[1:1000, ], w = 2000)
end <- Sys.time()

gh_time <-end-start

answer_gh <- gh_time * (n_gh/1000)

print(answer_gh)
```

To run greed_heuristic() with n = 1000000 takes answer_gh.