---
title: "Knapsack package"
author: "Henrik Karlsson and Milda Pocevicuite"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Knapsack package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Data used
The following example data is being used for all knapsack solutions in this package. 

Our package depends on dplyr.

```{r data, warning = FALSE, comment= NA, message=FALSE}
library(dplyr)

set.seed(42)
n <- 2000
knapsack_objects <-
    data.frame(
        w=sample(1:4000, size = n, replace = TRUE),
        v=runif(n = n, 0, 10000)
    )
```

# Brute Force Search


```{r}

powerset <- function(items) {
    len = length(items)
    l = vector(mode = "list", length = 2 ^ len)
    l[[1]] = numeric()
    counter = 1L
    for (x in 1L:length(items)) {
        for (subset in 1L:counter) {
            counter = counter + 1L
            l[[counter]] = c(l[[subset]], items[x])
        }
    }
    return(l)
}

brute_force_knapsack <- function(x, w, parallel = FALSE) {
    stopifnot(is.data.frame(x),
              apply(x, c(1, 2), is.numeric),
              is.numeric(w))
    
    stopifnot(x > 0,
              length(w) == 1,
              w > 0,
              is.logical(parallel))
    
    if (parallel == FALSE) {
        # initiate variables
        n <- nrow(x)
        best_v <- 0
        chosen_items <- c()
        items <- 1:n
        sets <- powerset(items)
        
        # loop through all possible sets (ommit empty set)
        for (i in 2:length(sets)) {
            c_sets <- unlist(sets[i])
            set_w <- 0
            set_v <- 0
            j <- 1
            # loop through the elements in the set
            while (j <= length(c_sets) && set_w <= w) {
                row <- c_sets[j]
                set_w <- set_w + x[row, 1]
                set_v <- set_v + x[row, 2]
                j <- j + 1
            }
            
            # compare the value of this set to the previous best value
            if (set_v > best_v && set_w <= w) {
                best_v <- round(set_v, 0)
                chosen_items <- c_sets
            }
        }
        result <- list(value = round(best_v, 2), elements = chosen_items)
        return(result)
        
    } else{
        
        cores <- parallel::detectCores()
        
        selection <- unlist(mclapply(1:nrow(x), FUN = function(y) {
                combn(rownames(x), y, paste, collapse = " ")
            }, mc.cores = cores
        ))
        
        sum_w <- unlist(mclapply(1:nrow(x), FUN = function(y) {
                (combn(x[, "w"], y, sum))
            }, mc.cores = cores
            ))
        
        sum_v <- unlist(mclapply(1:nrow(x), FUN = function(y) {
                (combn(x[, "v"], y, sum))
            }, mc.cores = cores
        ))
        
        max_value <- max(sum_v[which(sum_w < w)])
        max_weight <- max(sum_w[which(sum_w < w)])
        elements <- selection[which(sum_v == max_value & sum_w <= w)]
        
        return(list(
            value = round(max_value, 0),
            weight = max_weight,
            element = elements
        ))
        
    }
}

```

# Dynamic Programming


# Greedy Heuristic

The greedy_heuristic function takes two arguments, **x** which is the data frame containing weight and value and **w** which is the total maximum allowed weight of the function. The function computes the density $density = \frac{v}{w}$, sorts the data in in descending density order and pick out the objects with higest density within the the allowed maximum weight, w.

```{r, greedy_heuristic}
greedy_heuristic <- function(x = knapsack_objects, w){
    stopifnot(is.numeric(w))
    
    x <- x %>%
        mutate(element = seq_along((x[[1]])),
               density = v/w) %>% 
        arrange(desc(density))
    
    # Start values
    sum_weight <- 0
    sum_value <- 0
    weight_check <- 0
    selected_elements <- vector()
    i <- 1
    
    while (weight_check <= w) {
        sum_weight <- sum_weight + x[i, "w"]
        sum_value <- sum_value + x[i, "v"]
        
        selected_elements[i] <- x[i, "element"]
        
        weight_check <- sum_weight + x[i + 1, "w"]
        i <- i + 1
    }
    
    return(list(value = round(sum_value, 0),
                sum_weight = sum_weight,
                selected_elements = selected_elements))
}
```

### How much time does it take to run the algorithm?
```{r, gh_time}
n_gh <- 1000000

# Option 1
system.time(greedy_heuristic(x = knapsack_objects[1:1000, ], w = 2000))

# Option 2
start <- Sys.time()
greedy_heuristic(x = knapsack_objects[1:1000, ], w = 2000)
end <- Sys.time()
gh_time <- end-start
gh_time



answer_gh <- gh_time * (n_gh/1000)

print(answer_gh)
```

To run greed_heuristic() with n = 1000000 takes answer_gh.